---
title: "Simple example of basic PvSTATEM package pre-release version functionalities"
author: "Tymoteusz Kwieciński"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Simple example of basic PvSTATEM package pre-release version functionalities}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{ggplot2}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

# Introduction
The basic functionality of the `PvSTATEM` package is reading raw MBA data. To present the functionalities of the package, we created a simple, artificial `csv` file, which is pre-loaded into the package.

```{r}
library(PvSTATEM)

plate_filepath <- system.file("extdata", "artificial_plate.csv", package = "PvSTATEM", mustWork = TRUE) # get the filepath of the csv dataset


plate <- read_data(plate_filepath) # read the data with the layout
```

After the plate is successfully loaded, we can look into some basic information about the plate.

```{r}
plate$summary()

plate$summary(include_names = TRUE) # more detailed summary - names of the types

plate$get_sample(1)
plate$get_sample_by_type("STANDARD CURVE")[[1]]

plate$sample_names
plate$analyte_names
```


## Warnings
Our scripts are designed to catch potential errors in the data. If there are any warnings, they will be stored in the `warnings` field of the plate object. The warnings can be accessed by the `$warnings` field of the plate object or by the `$warnings` field of the sample object.

```{r}
## warnings ----------------------------------------------------------------

plate$warnings

plate$get_sample(5)$warnings

plate$get(analyte = "AMA1", sample = 1)

plate$get_sample_by_type("STANDARD CURVE")[[1]]$warnings
```



## simple standard curve plotting
```{r}
plot_standard_curve_antibody(plate, antibody_name = "AMA1")

plate$blank_adjustment()

plot_standard_curve_antibody(plate, antibody_name = "AMA1")
```
