---
title: "Basic SerolyzeR functionalities"
author: "Tymoteusz Kwieciński"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Basic SerolyzeR functionalities}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{ggplot2}
  %\VignetteDepends{nplr}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  dpi = 50,
  out.width = "70%"
)
```

## Reading the plate object


The basic functionality of the `SerolyzeR` package is reading raw MBA data.
To present the package's functionalities, we use a sample dataset from the Covid OISE study, which is pre-loaded into the package.
You might want to replace these variables with paths to your files on your local disk, to test the package on your data. 

Firstly, let us load the dataset as the [`Plate`] object.

```{r}
library(SerolyzeR)

plate_filepath <- system.file("extdata", "CovidOISExPONTENT.csv", package = "SerolyzeR", mustWork = TRUE) # get the filepath of the csv dataset

layout_filepath <- system.file("extdata", "CovidOISExPONTENT_layout.xlsx", package = "SerolyzeR", mustWork = TRUE)


plate <- read_luminex_data(plate_filepath, layout_filepath) # read the data

plate
```

The [`Plate`] is an object which encapsulates all the information about the plate, such as the sample names, analyte names, and the MFI values. For more details about the object, check the documentation by executing the command `?Plate`.

The `read_luminex_data` function used above, reads the data from the CSV plate file and the layout file and outputs the [`Plate`] object. This method allows also for more advanced options, such as overriding the information from the files, or setting the data reading options, e.g. the Luminex file format (`xPONENT` or `INTELLIFLEX`) or the file separators and data encoding. For more details about the function, we encourage you to explore the documentation by executing the command `?read_luminex_data`.

It also detects automatically the sample types, using an internal method [`translate_sample_names_to_sample_types`]. The method uses the sample names, sourced either from the layout file or the CSV plate file, to determine the sample types. The types of samples are then stored in the `sample_types` field of the plate object. The sample types are used to distinguish between the standard curve samples, control samples and the test samples. The documentation of this function contains more details about the sample types detection.


## Processing the whole plate

Once we have loaded the plate object, we may process it using the function [`process_plate`].
This function fits a model to each analyte using the standard curve samples. By default, it computes RAU values for each analyte using the corresponding model.
The computed RAU values are then saved to a CSV file in a specified folder with the parameter `output_dir` in a file with specified name `filename `, which by default is based on the plate name and the normalisation type - this function also allows normalisation for nMFI values, more details about this method may be found in the `nMFI` section of this document, or in documentation of `?get_nmfi` function.  

This function fits a model to each analyte using the standard curve samples. By default, it computes Relative Antibody Units (RAU) for each analyte based on the fitted model. The computed RAU values are saved to a CSV file in a directory specified by the `output_dir` parameter. The output file is named according to the `filename` parameter. If no filename is provided, the function automatically generates one based on the plate name and the selected normalisation type. 

Additionally, this function supports normalisation for normalised Median Fluorescence Intensity (nMFI) values. More details on this normalisation approach can be found in the `nMFI` section of this document or in the documentation of [`get_nmfi()`].

To get more information about the function, check `?process_plate`. 

Below, we execute the function on the plate object and save the output to a temporary directory.

```{r}
example_dir <- tempdir(check = TRUE) # create a temporary directory to store the output
df <- process_plate(plate, output_dir = example_dir)
colnames(df)
```

We can take a look at a slice of the produced dataframe (as not to overcrowd the article).
```{r}
df[1:5, 1:5]
```
To automate the processing the plates, we can also use [`process_file`] and [`process_dir`] methods. The former processes a single file, while the latter processes all the files in a directory.


[`process_file`] parses the plate file and saves the normalised output files to the specified directory. An example execution of the function is shown below.


```{r}

process_file(plate_filepath, layout_filepath, output_dir = example_dir, generate_report = TRUE)
```

It saves `RAU` and `nMFI` values to the `output_dir`, working similarly to the [`process_plate`] function. Additionally it has an option to produce a single plate report, which can be enabled with `generate_report = TRUE` parameter. The report is saved in the same directory as the output files, with the name based on the plate name. For more detailed description of the reports, we refer to the vignette \code{vignette("SerolyzeR HTML reports", package="SerolyzeR")}.

[`process_dir`] processes all the files in the specified directory. The description of this method can be found in the separate vignette vignette("Multiplate SerolyzeR functionalities").


## Quality control and normalisation details

Apart from the `process_plate` function, the package provides a set of methods allowing for more detailed and advanced quality control and normalisation of the data.

### Plate summary and details

After the plate is successfully loaded, we can look at some basic information about it.

```{r}
plate$summary()

plate$summary(include_names = TRUE) # more detailed summary

plate$sample_names[1:5] # print some of the sample names
plate$analyte_names[1:4] # print some of the analyte names
```

The summary can also be accessed using the built-in generic method `summary`.

```{r}
summary(plate)
```

### Quality control

The package can plot the RAU along the MFI values, allowing manual inspection of the standard curve.

```{r}
plot_standard_curve_analyte(plate, analyte_name = "OC43_S")
```

We can also plot the standard curve for different analytes and data types.
A list of all available analytes on the plate can be accessed using the command `plate$analyte_names`.

By default, all the operations are performed on the `Median` value of the samples; this option can be selected from the `data_type` parameter of the function.

```{r}
plot_standard_curve_analyte(plate, analyte_name = "RBD_wuhan", data_type = "Mean")
plot_standard_curve_analyte(plate, analyte_name = "RBD_wuhan", data_type = "Avg Net MFI")
```

This plot may be used to assess the standard curve's quality and anticipate some potential issues with the data.
For instance, if we plotted the standard curve for the analyte, `ME`, we could notice that the `Median` value of the sample with RAU of `39.06` is abnormally large, which may indicate a problem with the data.

```{r}
plot_standard_curve_analyte(plate, analyte_name = "ME")
plot_standard_curve_analyte(plate, analyte_name = "ME", log_scale = "all")
```

The plotting function has more options, such as selecting which axis the log scale should be applied or reversing the curve.
More detailed information can be found in the function documentation, accessed by executing the command `?plot_standard_curve_analyte`.

Another valuable method of inspecting the potential errors of the data is `plot_mfi_for_analyte`.
This method plots the MFI values of standard curve samples for a given analyte along the boxplot of the MFI values of the test samples.

It helps identify the outlier samples and check if the test samples are within the range of the standard curve samples.

```{r}
plot_mfi_for_analyte(plate, analyte_name = "OC43_S")

plot_mfi_for_analyte(plate, analyte_name = "Spike_6P")
```

For the `Spike_6P` analyte, the MFI values don't fall within the range of the standard curve samples, which could be problematic for the model.
The test RAU values will be extrapolated (up to a point) from the standard curve, which may lead to incorrect results.


The package allows also a straightforward way to adjust the MFI values using the background samples. To do so, we can use the `plate$blank_adjustment` method, as below:

```{r}
plate$blank_adjusted # verify if the data is already adjusted
```

```{r}
plate$blank_adjustment()
```

### Normalisation

After inspection, we may create the model for the standard curve of a specific antibody.
The model is fitted using the `nplr` package, which provides a simple interface for fitting n-parameter logistic regression models. Still, to create a more straightforward interface for the user, we encapsulated this model into our own class called `Model` for simplicity.
The detailed documentation of the `Model` class can be found by executing the command `?Model`.

The model is then used to predict the RAU values of the samples based on the MFI values.

#### RAU vs dilution

To distinguish between actual dilution values (the ones known for the standard curve samples) from the dilution predictions (obtained using the fitted standard curve), we introduced into our package a unit called RAU (Relative Antibody Unit) which is equal to the dilution **prediction** multiplied by a $1,000,000$ to provide a more readable value.

#### Inner nplr model

`nplr` package fits the model using the formula:

$$ y = B + \frac{T - B}{[1 + 10^{b \cdot (x_{mid} - x)}]^s},$$ where:

-   $y$ is the predicted value, MFI in our case,

-   $x$ is the independent variable, dilution of the standard curve samples in our case,

-   $B$ is the bottom plateau - the right horizontal asymptote,

-   $T$ is the top plateau - the left horizontal asymptote,

-   $b$ is the slope of the curve at the inflection point,

-   $x_{mid}$ is x-coordinate at the inflection point,

-   $s$ is the asymmetric coefficient.

This equation is referred to as the Richards' equation.
More information about the model can be found in the `nplr` package documentation.

#### Predicting RAU

By reversing that logistic function, we can predict the dilution of the samples based on the MFI values.
The RAU value is then the predicted dilution of the sample multiplied by $1,000,000$.

To limit the extrapolation error from above (values above maximum RAU value \eqn{\text{RAU}_{max}} for the standard curve samples), we clip all predictions above \eqn{M = RAU_{max} + \text{over\_max\_extrapolation}} to $M$ where `over_max_extrapolation` is user controlled parameter to the `predict` function.
By default `over_max_extrapolation` is set to $0$.

#### Usage

By default, the `nplr` model transforms the x values using the log10 function. To create a model for a specific analyte, we use the `create_standard_curve_model_analyte` function, which fits and returns the model for the analyte.

```{r}
model <- create_standard_curve_model_analyte(plate, analyte_name = "OC43_S")

model
```

Since our `model` object contains all the characteristics and parameters of the fitted regression model.
The model can be used to predict the RAU values of the samples based on the MFI values.
The output above shows the most critical parameters of the fitted model.

The predicted values may be used to plot the standard curve, which can be compared to the sample values.

```{r}
plot_standard_curve_analyte_with_model(plate, model, log_scale = c("all"))
plot_standard_curve_analyte_with_model(plate, model, log_scale = c("all"), plot_asymptote = FALSE)
```

Apart from the plotting, the package can predict the values of all the samples on the plate.

```{r}
mfi_values <- plate$data$Median$OC43_S
head(mfi_values)

predicted_rau <- predict(model, mfi_values)

head(predicted_rau)
```

The dataframe contains original MFI values and the predicted RAU values based on the model.

In order to allow extrapolation from above (up to a certain value) we can set `over_max_extrapolation` to a positive value.
To illustrate that we can look at prediction plots.
The `plot_standard_curve_analyte_with_model` takes any additional parameters and passes them to a `predict` method so we can visually see the effect of the `over_max_extrapolation` parameter.

```{r}
model <- create_standard_curve_model_analyte(plate, analyte_name = "Spike_6P")
plot_standard_curve_analyte_with_model(plate, model, log_scale = c("all"))
```

```{r}
plot_standard_curve_analyte_with_model(plate, model, log_scale = c("all"), over_max_extrapolation = 100000)
```

### nMFI
In some cases, the RAU values cannot be reliably calculated. This may happen when the MFI values of test samples are way higher than those of the standard curve samples. In that case, to avoid extrapolation but to be still able to compare the samples across the plates, we introduced a new unit called nMFI (Normalized MFI). The nMFI is calculated as the MFI value of the test sample divided by the MFI value of the standard curve sample with the selected dilution value.

nMFI values of the samples can be calculated in two ways - using the `get_nmfi` function or with the `process_plate` function that also saves the output into the CSV file by setting the `normalisation_type` parameter to `nMFI` in the `process_plate` function. By default the output will be saved as a file with the same name as the plate name but with the `_nMFI` suffix. 

```{r}
nmfi_values <- get_nmfi(plate)

# process plate with nMFI normalisation

df <- process_plate(plate, output_dir = example_dir, normalisation_type = "nMFI")
df[1:5, 1:5]
```

