---
title: "Multiplate SerolyzeR functionalities"
author: "Tymoteusz Kwieciński"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Multiplate SerolyzeR functionalities}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{ggplot2}
  %\VignetteDepends{nplr}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  markdown: 
    wrap: sentence
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  dpi = 50,
  out.width = "70%"
)
```

# Introduction

In this tutorial we are going to cover all the possible options of reading and writing data from multiple plates implemented in `SerolyzeR` package.

This tutorial will not be focused on a detailed analysis of the package functionalities that are designed to analyse a single plate at a time, as this is already covered in the \code{vignette("Basic SerolyzeR functionalities", package="SerolyzeR")}.

## Reading the data

Firstly, we need to locate our dataset.
The `SerolyzeR` package has a preloaded dataset, which can be found with a command given below:

```{r}
base_dir <- system.file("extdata", "multiplate_tutorial", package = "SerolyzeR", mustWork = TRUE) # get the filepath of the directory containing all the files

list.files(base_dir) # list all the files
```

As you can see, this directory contains multiple files - both raw output from the multiplex machine and the layout files, that help to identify the samples on the plate.
Each of the plate files, e.g.
`P1_SEROPED_PBT_62PLEX_PLATE1_03312023_RUN001.csv` contains its corresponding layout file, for instance `P1_SEROPED_PBT_62PLEX_PLATE1_03312023_RUN001_layout.xlsx` .

We will explain in detail how the directory should be structured in order to read the data properly in the section below.
Let us first read the directory.

## Reading a whole directory

To read the directory as above, we need to import the library `SerolyzeR` and run a single command `process_dir`. This function reads all the files in the directory and processes them. It has an option to generate 4 basic kind of outputs (similarly to the `process_file` function). It is:
- MFI (mean fluorescence intensity) - the raw input MFI values, with optional blank adjustment only
- nMFI (normalised mean fluorescence intensity) - the normalised values, with optional blank adjustment and normalisation to the reference sample
- RAU (relative antibody units) - the normalised values, with optional blank adjustment transformed to the relative antibody units
- report - a quality control report generated with [`generate_plate_report`] function

These output types can be selected with the `output_types` parameter, which is a vector of all outputs to be generated. By default this parameter equals to `c("RAU", "nMFI", "MFI")`.


Now, we will discuss how this method works and describe its most important parameters. 
By default, it saves all the outputs into the input directory, this can be changed with the `output_dir` parameter. In this tutorial, we will set the output directory to a temporary directory, which can be created with the `tempdir()` function.
We also specify the format of the input files, which is `xPONENT` in this case. The `format` parameter can be either set to `xPONENT`, `INTELLIFLEX` or `NULL`. In case the format is `NULL`, the function will try to infer the format based on the filename. For more details, we refer the reader to the documentation of the ` function.

One of the most useful features of the `process_dir` function is that it enables us to do all the preprocessing steps in one go. By setting `merge_outputs` parameter to `TRUE`, the method will output a single file with all the data from all the plates. This is very useful when we want to analyse the data from multiple plates at once.

```{r}

library(SerolyzeR)

output_dir = file.path(tempdir(), "multiplate-tutorial") # create a temporary directory to store the output data
R.utils::mkdirs(output_dir)


plates <- process_dir(base_dir, format = "xPONENT", output_types = c("RAU", "nMFI", "MFI"), output_dir = output_dir, merge_outputs = TRUE, return_plates = TRUE)

```
Remember, if you want to read all your files properly, you need to follow the naming convention of the files. The layout file should have the same name as the plate file, with the added suffix `_layout`. The detailed description of the convention, along with the list of all the possible arguments can be found in the documentation of the function, which can be accessed with `?process_dir` command.

Now, let us investigate the merged output of the `process_dir` function, which contains the data for all the samples and analytes, for each data type: MFI, nMFI and RAU.


For detailed description of the normalised output for a single plate, we refer the reader to the \code{vignette("Basic SerolyzeR functionalities", package="SerolyzeR")}. For the detailed description of the reports, we refer the reader to the \code{vignette("SerolyzeR HTML reports", package="SerolyzeR")}. If you are interested in specific plots genrerated by our package, for quality control of both single and multiple plates, we refer you to the vignette \code{vignette("SerolyzeR plots")}. 

In this tutorial, we will just denote that `process_dir` function also allows for automatic generation of both multiplate and singleplate reports, which can be enabled by setting `generate_reports` parameter to `TRUE`.


Let us investigate the output directory and the files that were created.

```{r}
list.files(output_dir)
```
Since, we have selected the `merge_outputs` parameter to be `TRUE`, we should see 3 files in the output directory: `merged_MFI_{current_timestamp}.csv`, `merged_nMFI_{current_timestamp}.csv` and `merged_RAU_{current_timestamp}.csv`. 

```{r}
merged_RAU_filepath <- list.files(output_dir, pattern = "RAU", full.names = TRUE) # get the exact path to the RAU file


RAU_data <- read.csv(merged_RAU_filepath)

```

This file is a standard CSV dataframe. Its rows correspond to the samples and the columns to the analytes. The first column is the plate name from the sample originates, the second one is the sample name. The remaining columns contain the analytes. 

```{r}
RAU_data[1:5, 1:5]

```



