---
title: "Quick introduction to plots created by our package"
author: "Mateusz Nizwantowski"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Quick introduction to plots created by our package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{ggplot2}
  %\VignetteDepends{nplr}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

# Introduction

The `PvSTATEM` package provides a variety of plots that can be used to visualize the Luminex data. In this vignette, we will show how to use them. To present the package's functionalities, we use a sample dataset from the Covid OISE study, which is pre-loaded into the package. Firsly, let us load the dataset as the `plate` object.

```{r}
library(PvSTATEM)
library(ggplot2)

plate_filepath <- system.file("extdata", "CovidOISExPONTENT.csv", package = "PvSTATEM", mustWork = TRUE) # get the filepath of the csv dataset

layout_filepath <- system.file("extdata", "CovidOISExPONTENT_layout.xlsx", package = "PvSTATEM", mustWork = TRUE)


plate <- read_luminex_data(plate_filepath, layout_filepath) # read the data

plate
```

In this vignette, we will omit some validation funcionality and focus on the plots. 
After the plate is successfully loaded, we should validate it by looking at some basic information about it using summary function.
But we can also 

