---
title: "Raw data parsers"
author: "Jakub Grzywaczewski"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Raw data parsers}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

# Introduction

Our package's main purpose is to read, perform quality control, and normalise raw MBA data. 
The entire package is made to be as user-friendly as possible so in most of your code you will read the data from 
xPONENT or INTELLIFLEX file using the `read_luminex_data` function and interact with the created `Plate` object.

Under the hood the `read_luminex_data` function uses a specific function to read a data from a given format 
and later standardizes this output to finally create a `Plate` object.

This article will go a little bit deeper into the details of the our data parsers and show you how to use them 
even outside the PvSTATEM package.

# Basic data loading

The simplest way of loading a file is to use the `read_luminex_data` function with default values.

```{r}
library(PvSTATEM)

plate_filepath <- system.file("extdata", "CovidOISExPONTENT.csv", package = "PvSTATEM", mustWork = TRUE)
layout_filepath <- system.file("extdata", "CovidOISExPONTENT_layout.xlsx", package = "PvSTATEM", mustWork = TRUE)
plate <- read_luminex_data(plate_filepath, layout_filepath)
summary(plate)
# display a sample of the dataframe
data.frame(plate)[c(1, 2), c(1, 2)]
```

The function itself has many parameters that can be used to customize the reading process.

For example we can change which datatype we want to find in the file. By default the datatype we are looking for
is the Median MFI value, but we can change it the Mean MFI value across beads.

```{r}
plate <- read_luminex_data(plate_filepath, layout_filepath, default_data_type = "Mean")
summary(plate)
# display a sample of the dataframe
data.frame(plate)[c(1, 2), c(1, 2)]
```

For the full list of parameters and their description please refer to the `read_luminex_data` documentation.

# PvSTATEM as a MBA reader

The `read_luminex_data` function enforces additional constrains on the raw MBA data such as the sample
names should follow a specific pattern in order to be correctly classified. If your data does not follow

our standards but you still want to use our parsers you can directly use the format specific functions here:
- `read_xponent_format`
- `read_intelliflex_format`

For example let us read the xPONENT file above using the `read_xponent_format`.

```{r}
output <- read_xponent_format(plate_filepath)
typeof(output)
names(output)
output[["ProgramMetadata"]]
names(output[["Results"]])
# sample of the data
output[["Results"]][["Median"]][c(1, 2), c(1, 2, 3)]
```

We can see now that the output of that function is a nested list containing the information parsed from the file.
As the structure of the output may be different across the formats this is not the recommended way to read the data
but the package is open enough to allow you to do so.

