---
title: "Quick overview of reports generated by PvSTATEM package"
author: "Mateusz Nizwantowski"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{Quick overview of reports generated by PvSTATEM package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
  %\VignetteDepends{ggplot2}
  %\VignetteDepends{nplr}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = FALSE,
  comment = "#>",
  warning = FALSE,
  message = FALSE,
  dpi = 50,
  out.width = "70%"
)
```

# Introduction

The `PvSTATEM` package provides for now one report that can be generated using the `generate_plate_report` function. 
It is a HTML report that contains a summary of the plate. It was optimized for size to be easily sent via email. 
In the future, we plan to add to the package report about Levey-Jennings plots. For now, we will focus on the plate summary report. 
To present this functionality, we use a sample dataset from the Covid OISE study, which is pre-loaded into the package. 
Firstly, let us load the dataset as the `plate` object.

```{r}
library(PvSTATEM)

plate_filepath <- system.file("extdata", "CovidOISExPONTENT.csv", package = "PvSTATEM", mustWork = TRUE)
layout_filepath <- system.file("extdata", "CovidOISExPONTENT_layout.xlsx", package = "PvSTATEM", mustWork = TRUE)
plate <- read_luminex_data(plate_filepath, layout_filepath)

plate
```

# Generating the report

To generate the report, we need to use the `generate_plate_report` function. There is one parameter that needs to be set, which is the `plate` object.
Generation of the report takes a few seconds, so please be patient. It can take up to a minute for a large plate.

Default report has 4 main sections:
- The most important information about the plate
- The layout of the plate
- The preview of standard curves
- Detailed information about analytes, this sections has tabs for each analyte, where the user can select the analyte of interest.

